DeepPicar:​ ​A​ ​Low-cost​ ​Deep​ ​Neural​ ​Network-based​ ​Autonomous​ ​Car

### 2017-10-04 paper/evaluation.tex
eval D multimodel revision

The Raspberry Pi 3 Model B platform used in DeepPicar equip a Boardcom
BCM2837 SoC, which has a quad-core 64bit ARM Cortex-A53 cluster,
running at up to 1.2GHz. The chip also includes Broadcom's Videocore IV
GPU, although we did not use the GPU in our evaluation due to the lack
of sofware support (Tensorflow is not compatible with the GPU).
For software, we use Ubuntu MATE 16.04, Tensorflow 1.1 and python
2.7. We disabled DVFS (dynamic voltage frequency scaling) and
configure the clock speed of each core at the maximum 1.2GHz.

@@

In this experiment, we launch multiple instances of DeepPicar's DNN
model at the same time and measure its impact to their inference
timings.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/perf_vs_modelcnt}
  \caption{Timing impact of co-scheduling multiple DNNs. 1x1: one DNN
    model using one CPU core; 4x1: four DNN models each using one CPU;
    1x2: one DNN model using two CPU cores; 2x2: two DNN models each
    using two CPU cores.} 
  \label{fig:perf-vs-multimodel}
\end{figure}

%% \begin{figure}[h]
%%   \centering
%%   \includegraphics[width=.5\textwidth]{figs/2ModelChart}
%%   \caption{ Change in inferencing time when 2 models are run concurrently. }
%% \end{figure}

Figure~\ref{fig:perf-vs-multimodel} shows the results. In the figure,
X-axis shows the system configuration: \#of DNN models x \#of CPU
cores\/DNN. For example, '4x1' means running four DNN models each of
which is assigned to run on one CPU core, whereas '2x2' means running
two DNN models, each of which is assigned to run on two CPU
cores. The Y-axis shows the average inference timing.
Two bars in the left show the impact of co-scheduling four DNN
models. Compared to executing a single DNN model on one CPU core
(1x1), when four DNN models are co-scheduled (4x1), each model
suffers an average inference time increase of approximately 15 ms,
$\sim$24\%. On the other hand, when two DNN models, each using two CPU
cores, are co-scheduled, the average inference timing is increased by
about 7 ms, or 10\%, comapred to the baseline of running one model
using two CPU cores.

@@

\begin{table}[h]
\centering
  \begin{tabular} {| l | l | l |}
  \hline
  \textbf{num models} & 1 & 2 \\ \hline
  \textbf{L1 refs} &  3.04E+10 &  3.04E+10 \\ \hline
  \textbf{L1 misses} & 4.78E+08 & 4.91E+08 \\ \hline
  \textbf{L1 miss \%} & 1.58 & 1.61 \\ \hline
  \textbf{L2 refs} & 3.31E+09 &  3.91E+09 \\ \hline
  \textbf{L2 misses} & 3.68E+08 & 4.62E+08 \\ \hline
  \textbf{L2 miss \%} & 11.12 & 10.88 \\ \hline
  \end{tabular}
  \caption{ Effect of 2 simulatneous models on cache accesses. }
\end{table}

The increase in inference times, however, was not the result of
increased cache misses due to additional accesses. For all models, the
number of L1 misses was always $\sim$1.6\% of all  references. In the
2 model test, L2 cache misses remained at $\sim$11\% of all
references, and each model in the 4 model test missed $\sim$13\% of
all L2 references.

\begin{figure}[h]
  \centering
  \includegraphics[width=.5\textwidth]{figs/4ModelChart}
  \caption{Change in inferencing time when 4 models are run simultaneously. }
\end{figure}

\begin{table}[h]
\centering
  \begin{tabular} {| l | l | l |}
  \hline
  \textbf{num models} & 1 & 4 \\  \hline
  \textbf{L1 refs} & 2.78E+10 & 2.79E+10 \\ \hline
  \textbf{L1 misses} & 4.36E+08 & 4.53E+08 \\ \hline
  \textbf{L1 miss \%} & 1.57 & 1.63 \\ \hline
  \textbf{L2 refs} & 2.83E+09 & 3.36E+09 \\ \hline 
  \textbf{L2 misses} & 3.59E+08 & 4.43E+08 \\ \hline
  \textbf{L2 miss \%} & 12.68 & 13.19 \\ 
\hline
  \end{tabular}
  \caption{ Effect of 4 concurrent models on cache accesses. }
\end{table}

In all multimodel tests, it was found that a greater number of models run simultaneously resulted in 
interference that led to a noticable increase in the average inferencing time that wasn't 
attributable to a change in cache misses. Instead, the overall time increases can most likely be 
traced to an increase in cache latency. Even if the models had a consistent number of cache hits, it 
is highly probable that the contention for the cache increased the access time for each model, 
consequently increasing the time it took for the models to execute their operations.

In terms of real-time performance under a WCET of 50 ms, the DeepPicar was unsatisfactory in all 
multimodel tests. On average, the 2 model test saw models miss their deadline by 6-8 ms, and the 4 
model test was worse as each model went over their deadlines by 27-29 ms. As a result, it can be 
ascertained that, if a WCET of 50 ms is required, DeepPicar would not be able to perform as necessary.

>>>>

The Raspberry Pi 3 Model B platform used in DeepPicar equip a Boardcom
BCM2837 SoC, which has a quad-core 64bit ARM Cortex-A53 cluster,
running at up to 1.2GHz. Each core has 16K private I\&D caches, and all
cores share a 512KB L2 cache.
The chip also includes Broadcom's Videocore IV
GPU, although we did not use the GPU in our evaluation due to the lack
of sofware support (Tensorflow is not compatible with the GPU).
For software, we use Ubuntu MATE 16.04, Tensorflow 1.1 and python
2.7. We disabled DVFS (dynamic voltage frequency scaling) and
configure the clock speed of each core at the maximum 1.2GHz.

@@

In this experiment, we launch multiple instances of DeepPicar's DNN
model at the same time and measure its impact to their inference
timings. In particular, we are interested in how shared resource
contention affect inference timing.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/perf_vs_modelcnt}
  \caption{Timing impact of co-scheduling multiple DNNs. 1x1: one DNN
    model using one CPU core; 4x1: four DNN models each using one CPU;
    1x2: one DNN model using two CPU cores; 2x2: two DNN models each
    using two CPU cores.} 
  \label{fig:perf-vs-modelcnt}
\end{figure}

%% \begin{figure}[h]
%%   \centering
%%   \includegraphics[width=.5\textwidth]{figs/2ModelChart}
%%   \caption{ Change in inferencing time when 2 models are run concurrently. }
%% \end{figure}
%% \begin{figure}[h]
%%   \centering
%%   \includegraphics[width=.5\textwidth]{figs/4ModelChart}
%%   \caption{Change in inferencing time when 4 models are run simultaneously. }
%% \end{figure}

Figure~\ref{fig:perf-vs-modelcnt} shows the results. In the figure,
X-axis shows the system configuration: \#of DNN models x \#of CPU
cores\/DNN. For example, '4x1' means running four DNN models each of
which is assigned to run on one CPU core, whereas '2x2' means running
two DNN models, each of which is assigned to run on two CPU
cores. The Y-axis shows the average inference timing.
Two bars in the left show the impact of co-scheduling four DNN
models. Compared to executing a single DNN model on one CPU core
(1x1), when four DNN models are co-scheduled (4x1), each model
suffers an average inference time increase of approximately 15 ms,
$\sim$24\%. On the other hand, when two DNN models, each using two CPU
cores, are co-scheduled, the average inference timing is increased by
about 7 ms, or 10\%, comapred to the baseline of running one model
using two CPU cores.

@@

%% \begin{table}[h]
%% \centering
%%   \begin{tabular} {| l | l | l |}
%%   \hline
%%   \textbf{num models} & 1 & 2 \\ \hline
%%   \textbf{L1 refs} &  3.04E+10 &  3.04E+10 \\ \hline
%%   \textbf{L1 misses} & 4.78E+08 & 4.91E+08 \\ \hline
%%   \textbf{L1 miss \%} & 1.58 & 1.61 \\ \hline
%%   \textbf{L2 refs} & 3.31E+09 &  3.91E+09 \\ \hline
%%   \textbf{L2 misses} & 3.68E+08 & 4.62E+08 \\ \hline
%%   \textbf{L2 miss \%} & 11.12 & 10.88 \\ \hline
%%   \end{tabular}
%%   \caption{ Effect of 2 simulatneous models on cache accesses. }
%% \end{table}

%% \begin{table}[h]
%% \centering
%%   \begin{tabular} {| l | l | l |}
%%   \hline
%%   \textbf{num models} & 1 & 4 \\  \hline
%%   \textbf{L1 refs} & 2.78E+10 & 2.79E+10 \\ \hline
%%   \textbf{L1 misses} & 4.36E+08 & 4.53E+08 \\ \hline
%%   \textbf{L1 miss \%} & 1.57 & 1.63 \\ \hline
%%   \textbf{L2 refs} & 2.83E+09 & 3.36E+09 \\ \hline 
%%   \textbf{L2 misses} & 3.59E+08 & 4.43E+08 \\ \hline
%%   \textbf{L2 miss \%} & 12.68 & 13.19 \\ 
%% \hline
%%   \end{tabular}
%%   \caption{ Effect of 4 concurrent models on cache accesses. }
%% \end{table}

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/l2missrate_vs_modelcnt}
  \caption{L2 cache miss rates of different neural network and core
    assignments. X-axis is the same as Figure~\ref{fig:perf-vs-modelcnt}.} 
  \label{fig:l2missrate-vs-modelcnt}
\end{figure}

These increases in inference times in the co-scheduled scenarios are
expected and likely caused by contention in the shared hardware
resources such as the shared L2 cache and/or the DRAM controller.
To further analyze the source of contention, we use hardwawre
performance counters of the processor. Specifically, we measure L2
miss rates of the DNN models first in isolation and then after
co-scheduling another models. If the shared L2 cache is the primary
source of inteference, then the measured L2 miss rates would
increase. Figure~\ref{fig:l2missrate-vs-modelcnt} shows the results.
As can be see in the figure, L2 miss rates are largely unchanged
regardless whether multiple models are co-scheduled or not. This
suggests that the shared L2 cache is not the bottleneck that caused
execution time increases. In other words, DNN models appear to be not
sensitive to the shared L2 cache space. % L2 partitioning is probably
                                % not going to be useful.
Instead, we hypothesize that it is likely caused by the memory
controller---the only other major shared hardware source---where
memory requests from different CPU cores contend, which would result
in increased memory access latency. While some Intel processors
provide uncore hardware counters that can measure average memory
access latency~\cite{ye2016maracas}, we were not able to identify
whether such hardware counters exist in the BCM2837 processor of
Raspberry Pi 3 due to the lack of documentation. Instead, in the next
experiment, we use memory intensive synthetic benchmark to test the
hypothesis.

%% The increase in inference times, however, was not the result of
%% increased cache misses due to additional accesses. For all models, the
%% number of L1 misses was always $\sim$1.6\% of all  references. In the
%% 2 model test, L2 cache misses remained at $\sim$11\% of all
%% references, and each model in the 4 model test missed $\sim$13\% of
%% all L2 references.

%% In all multimodel tests, it was found that a greater number of models
%% run simultaneously resulted in interference that led to a noticable
%% increase in the average inferencing time that wasn't attributable to a
%% change in cache misses. Instead, the overall time increases can most
%% likely be traced to an increase in cache latency. Even if the models
%% had a consistent number of cache hits, it is highly probable that the
%% contention for the cache increased the access time for each model,
%% consequently increasing the time it took for the models to execute
%% their operations.

%% In terms of real-time performance under a WCET of 50
%% ms, the DeepPicar was unsatisfactory in all multimodel tests. On
%% average, the 2 model test saw models miss their deadline by 6-8 ms,
%% and the 4 model test was worse as each model went over their deadlines
%% by 27-29 ms. As a result, it can be ascertained that, if a WCET of 50
%% ms is required, DeepPicar would not be able to perform as necessary.
